%%*****************************************************************************
%% $Id: errors.tex,v 1.1 1995/03/06 23:04:17 gerd Exp $
%%*****************************************************************************
%% Author: Gerd Neugebauer
%%-----------------------------------------------------------------------------


\section{\ProCom\ Messages}

\def\MSG#1{\item[#1]\ \\}
\def\MSGS#1#2{\item[#1]\vspace{-1ex}\item[#2]\ \\}

\begin{description}\itemsep=0pt
  \MSG{*** ProCom Error: The option prover is not an atom: {\em Prover}}
  The value of the option |prover| is of the form {\tt procom({\em Prover}\/)}
  and {\em Prover}\/ is not an atom. The atom should denote a \CaPrI{} module
  to be used for the compilation.

  This message is issued right after the matrix has been read, i.e. after the
  filters bt before the reductions are applied.

  \MSG{--- ProCom: CaPrI module {\em Prover}\/ loaded.}
  The dynamic \CaPrI{} module {\em Prover}\/ has been loaded into \ProTop.

  \MSG{--- ProCom prover {\em Prover}\/ is not applicable.}
  The prover named in the option |prover| has requirements which are not
  fulfilled. Thus \ProCom{} will try to find another prover which is
  applicable.

  \MSG{--- ProCom prover {\em Prover}\/ doesn't fulfill it's requirements.}
  The prover named in the option |prover| has requirements which are not
  fulfilled. Thus \ProCom{} will try to find another prover which is
  applicable.

  \MSG{--- ProCom: Using alternate prover {\em Prover}}
  A prover {\em Prover}\/ has been found which is applicable.

  \MSG{*** ProCom Error: No prover is applicable.}
  The attempt to find an applicable prover has failed. This means that the
  user has to name one or no prove can be started.

  \MSG{--- ProCom: Using dynamic module {\em Prover}}
  \ProCom{} uses a module which is not initially compiled into the system.

  \MSG{*** ProCom Error: Compilation aborted abnormally.}
  The compilation failed. This can either be caused by an error which
  identifies itself with some kind of message or due to a bug in \ProTop.

  \MSG{*** ProCom Error: Prover {\em Prover}\/ is not loaded.}
  The request to check a prover {\em Prover}\/ which is not loaded has been
  encountered.

  \MSG{*** ProCom Error: Unspecified descriptor encountered (var).}
  A descriptor contains a variable where an instruction is expected.

  \MSGS{*** ProCom. Initialization failed for search strategy {\em Search}}
       {--- Failure ignored. Trying to continue anyway.}
  The initialization procedure for a search strategy failed. This is usually a
  bug. The initialization should succeed always. \ProCom{} tries to continue
  and ignore the misbehaviour.

  \MSGS{*** ProCom. Invalid search strategy selected: {\em Search}}
       {--- Using iterative\_deepening(1,1,1) instead.}
  A search strategy has been specified in the option |search| which is not
  defined for \ProCom. \ProCom{} recovers this error by switching to the
  iterative deepening search strategy.

  \MSG{*** ProCom ignoring template with variable type: template({\em A},{\em
      B}\/)}
  A variable has been found for {\em A}.

  \MSG{*** ProCom Error: Template map has inconsistent length.}
  An incorrect use of the map directive has been encountered.

  \MSG{*** Library {\em Lib}\/ could not be found.}
  \MSG{*** Predicate {\em Spec}\/ could not be found.}
  \MSG{*** Library {\em File}\/ could not be found.}
  \MSG{*** Reorder module {\em Module}\/ is not loaded.}
  \MSG{*** Reorder predicate {\em Spec}\/ does not exist in module {\em
      Module}}
  \MSG{}
\end{description}



%
% Local Variables: 
% mode: latex
% TeX-master: "manual"
% End: 
